Constructing Biological Regulatory Networks from Process Hitting models

### 2013-07-26 main.tex
big application example + computation results on other examples
check TODOs in Introduction
finish proofreading of Examples
TODOs
cleanup
fixes after Morgan's review

\usepackage{prettyref}
\newrefformat{def}{Def.~\ref{#1}}
\newrefformat{fig}{Fig.~\ref{#1}}
\newrefformat{pro}{Property~\ref{#1}}
\newrefformat{pps}{Proposition~\ref{#1}}
\newrefformat{lem}{Lemma~\ref{#1}}
\newrefformat{thm}{Theorem~\ref{#1}}
\newrefformat{sec}{Sect.~\ref{#1}}
\newrefformat{ssec}{Subsect.~\ref{#1}}
\newrefformat{sssec}{Subsect.~\ref{#1}}
\newrefformat{suppl}{Appendix~\ref{#1}}
\newrefformat{eq}{Eq.~\eqref{#1}}
\newrefformat{ex}{Example~\ref{#1}}
\def\pref{\prettyref}

@@

% Commandes À FAIRE
%\usepackage{color} % Couleurs du texte
\newcommand{\warnul}[1]{\textcolor{red}{\underline{#1}}}
\newcommand{\rewriteng}[1]{\textcolor{blue}{#1}}
\newcommand{\rewriteil}[1]{\rewriteng{\textbf{[#1]}}}
\newcommand{\rewrite}[1]{\rewriteil{REWRITE: #1}}
\newcommand{\towrite}[1]{\rewriteil{TOWRITE: #1}}
\newcommand{\todo}[1]{\textcolor{red}{\textbf{[TODO: #1]}}}
\newcommand{\new}[1]{\textcolor{darkgreen}{\textbf{[$^\bigstar$New:} #1\textbf{$^\bigstar$]}}}
\def\updated#1{{\color{magenta}#1}}
%\newcommand{\new}[1]{#1}

% Id est
\newcommand{\ie}{i.e. }

@@

On the other hand, the qualitative modeling of BRNs has been widely addressed using Ren\'e Thomas'
formalism, leading to numerous theoretical work and practical tools to understand emerging behaviors.

This paper establishes formal links between these two formalisms, permitting to appreciate their
complementary features.
First, the inference of the Interaction Graph (IG) from a PH models allows a compact representation of
the regulations that are effective for the dynamics.
From this IG, well-known results can be applied to derive global dynamical properties for the
modelled BRN.
Second, the inference of all Thomas models that are compatible with a given PH is tackled.
As the PH allows partial specification of cooperation between components, this inference
emphasizes the ability of PH to deal with large BRNs with incomplete knowledge on
cooperations, where Thomas' approach fails because of the combinatorics of parameters.

The inference of corresponding Thomas' models is provided using Answer Set Programming,
which allows notably an efficient enumeration of (possibly numerous) compatible parametrizations.
\todo{application}
\end{abstract}

>>>>

\usepackage{prettyref}
\newrefformat{def}{Def.~\ref{#1}}
\newrefformat{fig}{Fig.~\ref{#1}}
\newrefformat{pro}{Property~\ref{#1}}
\newrefformat{pps}{Proposition~\ref{#1}}
\newrefformat{lem}{Lemma~\ref{#1}}
\newrefformat{thm}{Theorem~\ref{#1}}
\newrefformat{sec}{Sect.~\ref{#1}}
\newrefformat{ssec}{Subsect.~\ref{#1}}
\newrefformat{sssec}{Subsect.~\ref{#1}}
\newrefformat{suppl}{Appendix~\ref{#1}}
\newrefformat{eq}{Eq.~\eqref{#1}}
\newrefformat{ex}{Example~\ref{#1}}
\newrefformat{tb}{Table~\ref{#1}}
\def\pref{\prettyref}

@@

% Commandes À FAIRE
%\usepackage{color} % Couleurs du texte
%\newcommand{\warnul}[1]{\textcolor{red}{\underline{#1}}}
%\newcommand{\rewriteng}[1]{\textcolor{blue}{#1}}
%\newcommand{\rewriteil}[1]{\rewriteng{\textbf{[#1]}}}
%\newcommand{\rewrite}[1]{\rewriteil{REWRITE: #1}}
%\newcommand{\towrite}[1]{\rewriteil{TOWRITE: #1}}
\newcommand{\todo}[1]{\textcolor{red}{\textbf{[TODO: #1]}}}
%\newcommand{\new}[1]{\textcolor{darkgreen}{\textbf{[$^\bigstar$New:} #1\textbf{$^\bigstar$]}}}
%\def\updated#1{{\color{magenta}#1}}
%\newcommand{\updatedmax}[1]{{\color{violet}#1}}
%\newcommand{\new}[1]{#1}

% Id est
\newcommand{\ie}{i.e.~}

@@

On the other hand, the qualitative modeling of BRNs has been widely addressed using Ren\'e Thomas'
formalism, leading to numerous theoretical work and practical tools to understand emerging behaviors.
In addition to be able to use the existing results on this formalism,
obtaining the Thomas representation of a PH model allows the iteratively refine a system.

This paper establishes formal links between these two formalisms, permitting to appreciate their
complementary features.
First, the inference of the Interaction Graph (IG) from a PH model gives a compact representation of
the regulations that are effective for the dynamics.
From this IG, well-known results can be applied to derive global dynamical properties for the
modeled BRN.
Second, the inference of all Thomas models that are compatible with a given PH is tackled.
As the PH allows partial specification of cooperation between components, this inference
emphasizes the ability of PH to deal with large BRNs with incomplete knowledge on
cooperations, where Thomas' approach fails because of the combinatorics of parameters.

The inference of corresponding Thomas' models is provided using Answer Set Programming,
which allows notably an efficient enumeration of (possibly numerous) compatible parametrizations.
\end{abstract}
